---
title: "Exports and Imports Forecast"
author: "Fabián Villarreal"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<br />



#### **Exports and Imports Forecast with Univariate Time Series**

This R script makes a prediction of Ecuador's oil and non-oil exports and imports, using ARIMA and Holt-Winters models.

```{r libraries, echo = TRUE, results = 'hide', warning = FALSE, message = FALSE}
# Libraries
library(tidyverse)
library(readxl)
library(foreign)
library(ggfortify)
library(timeDate)
library(forecast)
library(xts)
library(urca)
library(tseries)
library(xlsx)

rm(list=ls())
```



#### **Data**
```{r}
# Data
wd <- 'C:/Users/fvill/Dropbox/BCE/Otros/Exportacion_Importacion Forecast'
setwd(wd)

df <- read_excel('export_import_petro_no_petro.xlsx')
```

```{r plot1, fig.width = 7, fig.height = 4}
# Oil Exports
export_petr <- ts(df[,'export_petr'], frequency = 12, start = c(1990,1), end = c(2021,12))
autoplot(export_petr) + ggtitle('Oil Exports') + ylab('Million USD FOB')
```

```{r plot2, fig.width = 7, fig.height = 4}
# Non-Oil Exports
export_no_petr <- ts(df[,'export_no_petr'], frequency = 12, start = c(1990,1), end = c(2021,12))
autoplot(export_no_petr) + ggtitle('Non-Oil Exports') + ylab('Million USD FOB')
```

```{r plot3, fig.width = 7, fig.height = 4}
# Oil Imports
import_petr <- ts(df[,'import_petr'], frequency = 12, start = c(1990,1), end = c(2021,12))
autoplot(import_petr) + ggtitle('Oil Imports') + ylab('Million USD FOB')
```

```{r plot4, fig.width = 7, fig.height = 4}
# Non-Oil Exports
import_no_petr <- ts(df[,'import_no_petr'], frequency = 12, start = c(1990,1), end = c(2021,12))
autoplot(import_no_petr) + ggtitle('Non-Oil Imports') + ylab('Million USD FOB')
```
<br />



#### **Unit Root Tests**
Determine the number of differences to calculate

```{r unit_root1, collapse = TRUE}
# Oil Exports
ndiffs(export_petr)
nsdiffs(export_petr)
export_petr %>% ur.kpss() %>% summary() # Not stationary
adf.test(export_petr) # Not stationary
``` 

```{r unit_root2, collapse = TRUE}
# Non-Oil Exports
ndiffs(export_no_petr) 
nsdiffs(export_no_petr)
export_no_petr %>% ur.kpss() %>% summary() # Not stationary
adf.test(export_no_petr) # Not stationary
``` 

```{r unit_root3, collapse = TRUE}
# Oil Imports
ndiffs(import_petr) 
nsdiffs(import_petr) 
import_petr %>% ur.kpss() %>% summary() # Not stationary
adf.test(import_petr) # Not stationary
``` 

```{r unit_root4, collapse = TRUE}
# Non-Oil Imports
ndiffs(import_no_petr)
nsdiffs(import_no_petr)
import_no_petr %>% ur.kpss() %>% summary() # Not stationary
adf.test(import_no_petr) # Not stationary
```
<br />



#### **ARIMA model estimation**
<br />

##### **Oil Exports**
```{r arima1, collapse = TRUE}
export_petr.arima1 <- auto.arima(export_petr, seasonal = FALSE, d = 1)
export_petr.arima2 <- auto.arima(export_petr, seasonal = FALSE, d = 1,
                        approximation = FALSE, 
                        stepwise = FALSE)
export_petr.arima3 <- auto.arima(export_petr, seasonal = TRUE, d = 1)
export_petr.arima4 <- auto.arima(export_petr, seasonal = TRUE, d = 1,
                        approximation = FALSE, 
                        stepwise = FALSE)

# Determine the best model using the AIC criterion
export_petr_m <- rep(0,4)
export_petr_m[1] <- export_petr.arima1$aic
export_petr_m[2] <- export_petr.arima2$aic
export_petr_m[3] <- export_petr.arima3$aic
export_petr_m[4] <- export_petr.arima4$aic
export_petr_m
which(export_petr_m == min(export_petr_m))
``` 
```{r arima1_res, fig.width = 7, fig.height = 4}
checkresiduals(export_petr.arima2)
export_petr.arima2

autoplot(forecast(export_petr.arima2)) + 
  ggtitle('Oil Exports ARIMA(3,1,2)') + 
  ylab('Million USD FOB')

write.xlsx(as.data.frame(forecast(export_petr.arima2)),
           'ep_arima.xlsx', 
           sheetName = 'ExporPetro')
```

##### **Non-Oil Exports**
```{r arima2, collapse = TRUE}
export_no_petr.arima1 <- auto.arima(export_no_petr, seasonal = FALSE, d = 1)
export_no_petr.arima2 <- auto.arima(export_no_petr, seasonal = FALSE, d = 1,
                                    approximation = FALSE,
                                    stepwise = FALSE)
export_no_petr.arima3 <- auto.arima(export_no_petr, seasonal = TRUE, d = 1)
export_no_petr.arima4 <- auto.arima(export_no_petr, seasonal = TRUE, d = 1,
                                    approximation = FALSE,
                                    stepwise = FALSE)

# Determine the best model using the AIC criterion
export_no_petr_m <- rep(0,4)
export_no_petr_m[1] <- export_no_petr.arima1$aic
export_no_petr_m[2] <- export_no_petr.arima2$aic
export_no_petr_m[3] <- export_no_petr.arima3$aic
export_no_petr_m[4] <- export_no_petr.arima4$aic
export_no_petr_m
which(export_no_petr_m == min(export_no_petr_m))
``` 
```{r arima2_res, fig.width = 7, fig.height = 4}
checkresiduals(export_no_petr.arima4) # Estimate with an AR 4

export_no_petr.fit1 <- Arima(export_no_petr, order = c(4,1,2), seasonal = c(2,0,0))
checkresiduals(export_no_petr.fit1)
export_no_petr.fit1

autoplot(forecast(export_no_petr.fit1)) + 
  ggtitle('Non-Oil Exports SARIMA(4,1,2)(2,0,0)[12]') + 
  ylab('Million USD FOB')

write.xlsx(as.data.frame(forecast(export_no_petr.fit1)),
           'enp_arima.xlsx', 
           sheetName = 'ExporNoPetro')
```

##### **Oil Imports**
```{r arima3, collapse = TRUE}
import_petr.arima1 <- auto.arima(import_petr, seasonal = FALSE, d = 1)
import_petr.arima2 <- auto.arima(import_petr, seasonal = FALSE, d = 1,
                                 approximation = FALSE,
                                 stepwise = FALSE)
import_petr.arima3 <- auto.arima(import_petr, seasonal = TRUE, d = 1)
import_petr.arima4 <- auto.arima(import_petr, seasonal = TRUE, d = 1,
                                 approximation = FALSE,
                                 stepwise = FALSE)

# Determine the best model using the AIC criterion
import_petr_m <- rep(0,4)
import_petr_m[1] <- import_petr.arima1$aic
import_petr_m[2] <- import_petr.arima2$aic
import_petr_m[3] <- import_petr.arima3$aic
import_petr_m[4] <- import_petr.arima4$aic
import_petr_m
which(import_petr_m == min(import_petr_m))
``` 
```{r arima3_res, fig.width = 7, fig.height = 4}
checkresiduals(import_petr.arima1) # Estimate with an AR 1

import_petr.fit1 <- Arima(import_petr, order = c(1,1,2))
checkresiduals(import_petr.fit1)
import_petr.fit1

autoplot(forecast(import_petr.fit1)) + 
  ggtitle('Oil Imports ARIMA(1,1,2)') + 
  ylab('Million USD FOB')

write.xlsx(as.data.frame(forecast(import_petr.fit1)),
           'ip_arima.xlsx', 
           sheetName = 'ImporPetro')
```

##### **Non-Oil Imports**
```{r arima4, collapse = TRUE}
import_no_petr.arima1 <- auto.arima(import_no_petr, seasonal = FALSE, d = 1)
import_no_petr.arima2 <- auto.arima(import_no_petr, seasonal = FALSE, d = 1,
                                    approximation = FALSE,
                                    stepwise = FALSE)
import_no_petr.arima3 <- auto.arima(import_no_petr, seasonal = TRUE, d = 1)
import_no_petr.arima4 <- auto.arima(import_no_petr, seasonal = TRUE, d = 1,
                                    approximation = FALSE,
                                    stepwise = FALSE)

# Determine the best model using the AIC criterion
import_no_petr_m <- rep(0,4)
import_no_petr_m[1] <- import_no_petr.arima1$aic
import_no_petr_m[2] <- import_no_petr.arima2$aic
import_no_petr_m[3] <- import_no_petr.arima3$aic
import_no_petr_m[4] <- import_no_petr.arima4$aic
import_no_petr_m
which(import_no_petr_m == min(import_no_petr_m))
``` 
```{r arima4_res, fig.width = 7, fig.height = 4}
checkresiduals(import_no_petr.arima4)
import_no_petr.arima4

autoplot(forecast(import_no_petr.arima4)) + 
  ggtitle('Non-Oil Imports SARIMA(3,1,0)(2,0,0)[12]') + 
  ylab('Million USD FOB')

write.xlsx(as.data.frame(forecast(import_no_petr.arima4)),
           'inp_arima.xlsx', 
           sheetName = 'ImporNoPetro')
```
<br />



#### **Holt-Winters model estimation**
<br />

##### **Oil Exports**
```{r hw1_res, fig.width = 7, fig.height = 4}
export_petr_HW <- HoltWinters(export_petr)

autoplot(forecast(export_petr_HW)) + 
  ggtitle('Oil Exports HW') + 
  ylab('Million USD FOB')

write.xlsx(as.data.frame(forecast(export_petr_HW)),
           'ep_HW.xlsx', 
           sheetName = 'ExporPetro')
```

##### **Non-Oil Exports**
```{r hw2_res, fig.width = 7, fig.height = 4}
export_no_petr_HW <- HoltWinters(export_no_petr)

autoplot(forecast(export_no_petr_HW)) + 
  ggtitle('Non-Oil Exports HW') + 
  ylab('Million USD FOB')

write.xlsx(as.data.frame(forecast(export_no_petr_HW)),
           'enp_HW.xlsx', 
           sheetName = 'ExporNoPetro')
```

##### **Oil Imports**
```{r hw3_res, fig.width = 7, fig.height = 4}
import_petr_HW <- HoltWinters(import_petr)

autoplot(forecast(import_petr_HW)) + 
  ggtitle('Oil Imports HW') + 
  ylab('Million USD FOB')

write.xlsx(as.data.frame(forecast(import_petr_HW)),
           'ip_HW.xlsx', 
           sheetName = 'ImporPetro')
```

##### **Non-Oil Imports**
```{r hw4_res, fig.width = 7, fig.height = 4}
import_no_petr_HW <- HoltWinters(import_no_petr)

autoplot(forecast(import_no_petr_HW)) + 
  ggtitle('Non-Oil Imports HW') + 
  ylab('Million USD FOB')

write.xlsx(as.data.frame(forecast(import_no_petr_HW)),
           'inp_HW.xlsx', 
           sheetName = 'ImporNoPetro')
```
